---
title: "IC13: Predicting economic mobility"
author: "Solution"
date: "28 April 2020"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
opts_chunk$set(echo=FALSE, fig.align='center', fig.width=8, fig.height=8,
               cache=TRUE, autodep=TRUE, cache.comments=FALSE,
               message=FALSE, warning=FALSE)
```

Your goal is to use this data to predict economic mobility. Note that there are generally more observations than predictors.

```{r load-data}
mob = read.csv('mobility.csv')
```

1. Using `glmnet`, estimate 4 models: the linear model, ridge regression, the lasso, and the elastic net ($\alpha=.5$). Don't use the variables `ID`, `Name`, or `State` (Why?)

2. Plot the CV curves for each of the three regularized models (easy).

3. Use `lambda.min` to get a particular model for each of the regularized ones.

4. Plot the coefficients for each of the 4 models on one figure. What do you notice? Which features are most important?


```{r model-fitting}
library(glmnet)
linmod = lm(Mobility~.-ID-Name-State, data=mob, y=TRUE)
X = model.matrix(linmod)[,-1]
y = linmod$y
lasso = cv.glmnet(X, y)
ridge = cv.glmnet(X, y, alpha=0, lambda.min.ratio=1e-6)
enet = cv.glmnet(X,y,alpha=.5)
```

```{r cvplots}
par(mfrow=c(2,2))
plot(lasso)
plot(ridge)
plot(enet)
par(mfrow=c(1,1))
```


For `enet` and `lasso`, `lambda.1se` gives sparser models. For `ridge`, use `lambda.min` (more like GCV).

```{r models}
lasso1 = as.numeric(coef(lasso, 'lambda.min'))
enet1 = as.numeric(coef(enet, 'lambda.min'))
ridge1 = as.numeric(coef(ridge, 'lambda.min'))
```


```{r plot-coefs}
ord = order(coef(linmod))
df = data.frame(lm = coef(linmod)[ord], lasso = lasso1[ord],
                elnet = enet1[ord], ridge = ridge1[ord])
df$var = rownames(df) 
df %>% mutate(var = str_replace_all(var, "_"," ")) %>%
  pivot_longer(names_to='method',values_to ='estimate', -var) %>%
  ggplot(aes(y=var,x=estimate,color=method)) + geom_point() + 
  geom_vline(xintercept = 0) +
  cowplot::theme_cowplot() + 
  scale_color_viridis_d(direction = -1) +
  theme(axis.title.y = element_blank())
  #scale_x_continuous(trans = scales::pseudo_log_trans(sigma=.1))
```
