---
title: "In-class 3: R-squared"
author: "[Your names here]"
date: "30 January 2020"
output: 
  html_document:
    highlight: tango
---

## Instructions

In HW1 you wrote the following function to generate data from a linear model (see my solutions):

```{r functions, echo=TRUE}
generate.data <- function(n, p, sig.epsilon=1){
  ## This function takes 3 inputs (2 mandatory, 1 optional)
  ## n - the number of observations
  ## p - the number of predictors
  ## sig.epsilon - (optional) the sd of the normal noise (default=1 if omitted)
  X = matrix(rnorm(p*n), ncol=p) ## a matrix of standard normal RVs (n x p)
  epsilon = rnorm(n, sd = sig.epsilon) ## noise ~ N(0, sig.epsilon)
  beta = p:1 # p beta coefficients
  beta.0 = 3 # an intercept
  y = beta.0 + X %*% beta + epsilon # the linear model
  df = data.frame(y, X) # put it all in a data frame
  return(df) # output
}
```

You should complete the code below to see what happens to $R^2$ as you add additional _irrelevant_ predictors. Note that some lines have comments requiring you to add code. Others ask you questions which you should answer in the comments.

## Complete the following

```{r adding-vars, echo=TRUE, fig.width=4, fig.height=4}
library(tidyverse)
p = 3 # the number of true predictors
n = 100 # the total number of observations
pmax = 10 # the maximum number of predictors to examine, feel free to change this
stopifnot(pmax<n, pmax>p) # what does this do?
true.model = # add code here
extra.predictors = data.frame(matrix(rnorm(n*(pmax-p)), nrow=n)) 
names(extra.predictors) = paste0('X',(p+1):pmax) # why am I naming the extra predictors? What were their names before
full.data = bind_cols(true.model,extra.predictors)
rsq = double(pmax) # What do we call this step? We discussed this last week.
for(iter in 1:pmax){
  form = formula(paste('y ~ ', paste(names(full.data)[2:(iter+1)], collapse=' + '))) # annoying
  fit.mdl = lm(form, data=full.data)
  rsq[iter] = summary(fit.mdl)$r.sq # note how easy it is to get this
}
ggplot(data.frame(nvars=1:pmax, rsq),aes(x=nvars,y=rsq)) + 
  geom_path() + geom_point(color='red') + theme_minimal(base_family = 'serif')
```

## Questions to answer:

1. Does $R^2$ decrease? 
2. Why or why not?