---
title: "In class 4"
author: "[Your names here]"
date: "4 February 2020"
output: 
  html_document:
    highlight: tango
---

## Instructions

1. Rename this document with your student ID (not the 10-digit number, your IU username, e.g. `dajmcdon`). Include your buddy if you are working together.
1. Examine the function below called `generate.data`.
1. Answer the question about the marked line by `X`.
1. Fill in the missing bits in the `intervals` function.
1. In part 3, call the function `intervals` for 25 different correlations. Each time, save the output.
1. Repeat the above step 50 times. Find the average width for each correlation over the 50 replications.
1. Plot the average __width__ of the CI against `cors`.


## Functions to examine and complete

```{r functions-to-create}
library(ggplot2)
# This function generates Y, X1, and X2
## Inputs: n, the number of data points
##         cor, the correlation between X1 and X2
## Outputs: a data frame
generate.data <- function(n, cor){
  epsilon = rnorm(n)
  X = matrix(rnorm(2*n),nrow=n) %*% 
    chol(matrix(c(1, cor, cor,1), 2, 2))
  ## See the question below about X
  beta = c(3, 2, 1)
  Y = cbind(1, X) %*% beta + epsilon
  df = data.frame(Y, X)
  return(df)
}

## This function finds confidence interval widths 
## for the Linear model
## Inputs: n, the number of data points
##         cor, the correlation between X1 and X2
## Outputs: avg, the average width of the confidence 
##               intervals for the regression
intervals <- function(n, cor){
  df = generate.data(n, cor) 
  mdl = lm(Y~X1+X2, data=df)
  itvals =  ## Get the confidences intervals for the bhats (95%)
  widths =   ## Find the width of each interval
  avg =   ## get the average width, ignore intercept
  return(avg)
}

n = 250
ncors = 25
cors = 1 / (1 + exp(-seq(-5,5,length.out = ncors)))*2 - 1
```

## What is the the marked line in `generate.data` doing? Why did I multiply by some matrix?


## Running the simulation and plotting

```{r eval-and-plot}
## calculate the average width for each value of cors
## repeat the experiment 50 times
    
avg.widths = 

## plot cors (x-axis) vs. average widths (y-axis)

##    label your axes nicely
##    make the y-axis on the log scale
```

## How does correlation affect confidence intervals?

