---
title: "In-class 5"
author: "DJM"
date: "13 February 2020"
output: html_document
---

## Instructions

1. Rename this document with your student ID (not the 10-digit number, your IU username, e.g. `dajmcdon`). Include your buddy in the author field if you are working together.
1. I have given you code to generate data and fit 4 different models to the data. You should run through the code line by line in the console.
1. Discuss the questions with your neighbors. Write down answers.

## Generate data and fit models

```{r generate-data-and-fit}
generate.data = function(n, p=3){
  X = 5 + matrix(rnorm(3*n), n)
  beta = c(runif(p+1, -1,1))
  epsilon = rnorm(n)
  Y = exp(beta[1] + X %*% beta[-1] + epsilon) ## NOTE THIS LINE!!
  data.frame(Y,X)
}
set.seed(20200213)
n = 250
dat = generate.data(n)
formulae = lapply(
  c('Y~.', 
    'log(Y)~.',
    paste0('Y ~', paste(paste0('log(X',1:3,')'),collapse='+')),
    paste0('log(Y) ~', paste(paste0('log(X',1:3,')'),collapse='+'))), 
  as.formula)
all.the.models = lapply(formulae, function(x) lm(x, data=dat))
```

## Make QQ plots

```{r qq-plots}
## Base R version
#par(mfrow = c(2,2))
#for(i in 1:4){
#  qqnorm(residuals(all.the.models[[i]]))
#  qqline(residuals(all.the.models[[i]]))
#}
library(tidyverse)
resids = as_tibble(
  sapply(all.the.models, residuals), .name_repair = ~paste0("model",1:4))
resids %>% pivot_longer(everything()) %>%
  ggplot(aes(sample=value)) + geom_qq() + geom_qq_line() + 
  facet_wrap(~name, 2, scales = 'free_y')
```

## Calculate CV

```{r cv}
cv.lm = function(mdl) mean(residuals(mdl)^2 / (1-hatvalues(mdl))^2)
sapply(all.the.models, cv.lm)
```

## Questions to answer

1. Which of the 4 models is the correct one?

2. What do you notice in the Q-Q plots? Which ones look ok? Why?

3. Examine the hatvalues for the 4 different models. What do you notice?

4. Consider models 1 and 2. In these two cases, what is `residuals(mdl)` doing? Think about how the log transformation affects these two things.

5. Is it reasonable to compare the CV values for models 1 and 3 with those of models 2 and 4? Why or why not?

6. How should we decide which model to use? Note: This is a subtle issue without a correct answer in light of the previous question.