---
title: 'In-class 7: Bootstrapping'
author: "[Your name here]"
date: "27 February 2020"
output: html_document
---

## Instructions

1. Rename this document with your student ID (not the 10-digit number, your IU username, e.g. `dajmcdon`). Include your buddy in the author field if you are working together.
1. I have given you code to generate data and fit a linear model.
1. Follow the instructions to replicate the bootstrap from the text/lecture on this new data.
1. Answer the question at the end.


## Generate data and fit models

```{r generate-data-and-fit}
## Note: this is the same data as in Lecture 8.
trueFunction <- function(x) sin(x) + 1/sqrt(x) + 3
set.seed(1234)
n = 100
x = 1:n/n*2*pi
wiggly = data.frame(x = x,
                 y = trueFunction(x) + rnorm(n, 0, .75))
library(ggplot2)
ggplot(wiggly, aes(x, y)) + geom_point() + xlim(0,2*pi) + ylim(0,max(wiggly$y)) +
  stat_function(fun=trueFunction, color='red')
```

```{r bootstrap-code-from-chapter, echo=FALSE}
resample <- function(x) {
    sample(x, replace = TRUE)
}
resample.data.frame <- function(data) {
    sample.rows <- resample(1:nrow(data))
    return(data[sample.rows, ])
}
rboot <- function(statistic, simulator, B) {
    tboots <- replicate(B, statistic(simulator()))
    if (is.null(dim(tboots))) {
        tboots <- array(tboots, dim = c(1, B))
    }
    return(tboots)
}
bootstrap <- function(tboots, summarizer, ...) {
    summaries <- apply(tboots, 1, summarizer, ...)
    return(t(summaries))
}
equitails <- function(x, alpha) {
    lower <- quantile(x, alpha/2)
    upper <- quantile(x, 1 - alpha/2)
    return(c(lower, upper))
}
bootstrap.ci <- function(statistic = NULL, simulator = NULL, tboots = NULL,
    B = if (!is.null(tboots)) {
        ncol(tboots)
    }, t.hat, level) {
    if (is.null(tboots)) {
        stopifnot(!is.null(statistic))
        stopifnot(!is.null(simulator))
        stopifnot(!is.null(B))
        tboots <- rboot(statistic, simulator, B)
    }
    alpha <- 1 - level
    intervals <- bootstrap(tboots, summarizer = equitails, alpha = alpha)
    upper <- t.hat + (t.hat - intervals[, 1])
    lower <- t.hat + (t.hat - intervals[, 2])
    CIs <- cbind(lower = lower, upper = upper)
    return(CIs)
}
```

## Function 1

Write a function that estimates the linear model of $y$ on $x$ for any data set. Call your function `wiggly.estimator.lm`. It should return the slope and intercept. Run `lm` on the wiggly data (call the fitted model `wiggly.mdl`). Compare it's estimates to those of `wiggly.estimator.lm`.

```{r estimator-fcn}
wiggly.estimator.lm <- function(newdata) 
wiggly.mdl = 
coef(wiggly.mdl)
wiggly.estimator.lm(wiggly)
```

## Function 2

Write a function that resamples the residuals. Call it `wiggly.resids.resamp`.

```{r resid-sampler-fcn}
wiggly.resids.resamp <- function(){
}
```

## Function 3

Write a function to resample the data frame.

```{r data-frame-sampler}
new.wiggly.resamp <- function() {
}
```

## Use your functions

Use both samplers to get model-based and nonparametric confidence intervals for your linear model.

```{r bootstrap}
mbb.wiggly = wiggly.resids.resamp(wiggly)
npb.wiggly = 
mbb.wiggly
npb.wiggly
```

## Question to answer

Which bootstrap is more appropriate to use in this case? Why?